---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r include=FALSE}
knitr :: opts_chunk$set(echo = TRUE, cache = TRUE)
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r preprocessing}
filename1 = "activity.zip"
unzip(filename1)
activity <- read.csv(file = "activity.csv", header = TRUE, na.strings = "NA")
activity$date = as.Date(activity$date)
head(activity)
```



## What is mean total number of steps taken per day?
```{r total steps per day}
activity<- tbl_df(activity)

tot.steps.day<-activity%>%
  group_by(date)%>%
  summarize(total.steps = sum(steps, na.rm = TRUE))

#head(tot.steps.day)


##histogram Plot

hist_plot<- ggplot(tot.steps.day)+
  geom_histogram(aes(total.steps), bins = 15 , col ="blue", alpha = 1/2, fill = "red")+
  labs(x ="Number of steps", y ="Frequency", title = "Number of steps over two Year per Day")

print(hist_plot)

summary.tot.steps.day<-summary(tot.steps.day$total.steps)

mean.steps <- as.character(round(summary.tot.steps.day[["Mean"]],3))
median.steps<-as.character(round(summary.tot.steps.day[["Median"]],3))

```


*The mean and Median of total number of steps taken per day `r mean.steps` and `r median.steps` respectively*

## What is the average daily activity pattern?
```{r averageDailyPattern}
tot.steps.int<-activity%>%
  group_by(interval)%>%
  summarise(total.steps = mean(steps, na.rm = TRUE))

#time-series plot 
time_plot<- ggplot(tot.steps.int, aes(interval,total.steps))+
  geom_line( col = "blue")+
  labs(x =" 5-minute interval", y ="Number of steps", title = "Number of steps over years per interval")


print(time_plot)

time_interval <- tot.steps.int[which.max(tot.steps.int$total.steps),]
time_interval<-time_interval$interval

```

The time interval which have maximum steps across the days is `r time_interval`.

## Imputing missing values
```{r imputing NA}
activity.NA <- activity[which(is.na(activity$steps)),]
activity.NoNa<- activity[which(!is.na(activity$steps)),]

average.day <- activity.NoNa%>%
  group_by(date)%>%
  summarise(average.steps = mean(steps, na.rm = TRUE))

average.day$average.steps = as.integer(average.day$average.steps)
#head(average.day)

activity.NA$steps <- ifelse(activity.NA$date == average.day$date , 
                       average.day$average.steps,activity.NA$steps)

comb_activity<- rbind(activity.NoNa,activity.NA)

#head(comb_activity)

tot.steps.day.1<-comb_activity%>%
  group_by(date)%>%
  summarize(total.steps = sum(steps, na.rm = TRUE))

#head(tot.steps.day)


##histogram Plot

hist_plot<- ggplot(tot.steps.day.1)+
  geom_histogram(aes(total.steps), bins = 15 , col ="blue", alpha = 1/2, fill = "red")+
  labs(x ="Number of steps", y ="Frequency", title = "Number of steps over two Year per Day")

print(hist_plot)

summary.tot.steps.day.1<-summary(tot.steps.day.1$total.steps)

mean.steps.1 <- as.character(round(summary.tot.steps.day.1[["Mean"]],3))
median.steps.1<-as.character(round(summary.tot.steps.day.1[["Median"]],3))

```

*The mean and Median of total number of steps taken per day without imputing missing values `r mean.steps` and `r median.steps` respectively.*

*The mean and Median of total number of steps taken per day after imputing missig values `r mean.steps.1` and `r median.steps.1` respectively*

***So there is No change in mean and median with or without imputing values because we have filled the missing values with average of other values.***


## Are there differences in activity patterns between weekdays and weekends?

```{r activitypatterns weekdays and weekends}

activity.days <- comb_activity%>%
  mutate(day = weekdays(as.Date(activity$date)))%>%
  mutate(day = if_else(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday","Friday"),"weekday","weekend"))%>%
  group_by(interval,day)%>%
  summarise(average.steps = mean(steps, na.rm = TRUE))

#table(activity.days$day)

#time-series_plot

time_days_plot<- ggplot(activity.days, aes(interval,average.steps))+
  geom_line(col = "blue")+
  facet_grid(day~.)+
  labs(x = "Interval", y= "Average number of steps", title = "Average Number of steps across years")
  
  
print(time_days_plot)
  
```